---
title: "Project"
author: "Lucio Rezende Queiroz"
date: "March 22, 2017"
output:
  pdf_document: default
  html_notebook: default
  ioslides_presentation: default
  html_document: default
  word_document: default
---

# Viral Classification

# Setting envinroment

```{r}
args = commandArgs(trailingOnly=TRUE)
system('git config --global user.name "luciorq"')
system('git config --global user.email "luciorqueiroz@gmail.com"')
```

# Analysis

## Loading libraries

```{r}
#source("src/functions.R")
source("lib/rmd2rscript.R")
require(dplyr)
```

## Initializing varaibles

```{r}
sequences = "raw/contigs_MCV.fasta.txt"
db_name = "virus_ORF"
db_file = paste0("data/",db_name,".db")
```

# Extracting fasta files and creating database
```{r}
system2("bash",args = c("src/RunAllFasta.sh", sequences, db_name))
```

## Plotting Correlation

```{r}
system2("Rscript", args = c("--no-save", "src/PearsonCorrelation.R", db_file))
```

# Converting this notebook to an executable script

```{r}
rmd2rscript("dsRBP.Rmd")
```


# How can we improve?

- Ramdom Forests
- Permutation tests for clustering
- Use confidence interval for the Odds Ratio
- Have a good training database


## Acknowledgement

